---
title: "re_tidy"
author: "Leano Meier"
date: "2023-03-06"
output: html_document
---

#Perform the following steps:

1. Manually clean the data in the tab “Database S1” and save it as a CSV file that can be read into R.

2. In R, aggregate data per experiment and calculate the log-response ratio within each experiment.

3. Aggregate data across all experiments for different years since the start of the experiment, distinguishing    an early phase (<3 years since start), a mid-phase (3-6 years since start), and a late phase (>6 years since    start). Calculate the log-response ratio for each phase.

#Data input
```{r}
library(tidyverse)
#read dataset with read_csv2(), because of semikolon seperator
dataset1 <- readr::read_csv2("~/Desktop/Uni Bern/AGDS1/agds_report_leanomeier/data/1249534s1-s6.csv")
```

#calculate the log response ratio per experiment
```{r}
#calculate log response rate per datapoint
dataset1$log_response_rate <- log(dataset1$mean_increased_CO2/dataset1$mean_ambient_CO2)
#calculate mean log response rate per experiment
log_response_per_experiment <- dataset1 %>%
  group_by(dataset1$Experiment) %>% 
  summarise(log_response_rate = mean(dataset1$log_response_rate))
```

#Aggregate data across all experiments for different phases
```{r}
#making phases
dataset1$phase <- 0
for(i in 1:length(dataset1$Durration_years)){
  if(dataset1$Durration_years[i] < 3){
  dataset1$phase[i] <- 1
  }else if(dataset1$Durration_years[i] >=3 && dataset1$Durration_years[i] <= 6){
  dataset1$phase[i] <- 2
  }else if(dataset1$Durration_years[i] >6){
  dataset1$phase[i] <- 3
  }
}
#log response per phase
log_response_per_phase <- dataset1 %>%
  group_by(dataset1$phase) %>%
  summarise(log_response_rate_phase = mean(dataset1$log_response_rate))
```

